Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-05-26 WSSSPE2_report/WSSSPE2_review.tex
mostly changes in 7.1.3 - now done with 7

Subgroup A consisted of ten participants who discussed papers surrounding training
and successful community software initiatives.
Subgroup B discussed four papers and had six participants. Because of the nature
of the papers, training emerged as a common theme. However the conversation was
wide-ranging, including incentives, reproducibility, and funding to promote
sustainability.

@@

Three main actions were identified by Subgroup A that would be of interest to participants and that
would benefit the research community.  These focused on the desire to amplify the 
current community efforts in training and software engineering by conecting the current training initiatives.
( e.g., software/data carpentry, rOpenSci, OSS) of different durations at appropriate stages. Generally it 
was felt that a modicum of interaction between short term workshops (SwC, DC), medium term trainings (e.g., OSS),
and longer-term courses (e.g., BIDS Data Science) would benefit from coordinating curriculum,
discussing and aligning prerequisites, and coordinating timing of courses.  A training
coordination effort would go a long ways towards amplifying the value of the individual
efforts and make them all more effective.

\begin{itemize}
    \item Action 1: Create a roadmap of research software training initiatives. Such a roadmap would
        provide a taxonomy of training opportunities: what they deliver, and what is needed to know 
        going in (prerequisites). It would also show which recommended roadmap actions would have a clear 
        and immediate payoff, and which will have long-term payoffs.  The training roadmap would emphasize 
        time savings and efficiency gains to be had from each training.

    \item Action 2: Build a report card characterizing use of best practices in sceintific software. Generally,
        people felt that researchers would be very willing to migrate practices if they could identify
        where they needed to improve. Such a report card would ask simple questions to charactize use of best
        practices in science software, and could be structured similarly to Joel Spolsky's 'Software Maturity'
        questionairre (http://www.joelonsoftware.com/articles/fog0000000043.html).  The survey would create a 
        report card that shows areas where a project could improve, and then link those areas to specific 
        training offerings from the training taxonomy from Action 1.

    \item Action 3: Create science software review forum.  It was generally acknowledged that a little code review
        can have a tremendous impact on the quality of software in a project, but that sites for science code
        review are lacking.  While people can ask questions on the Stack Exchange sites, they are generally not open
        to questions of style, approach, or appropriateness, as they try to avoid subjective commentary.   Instead, 
        we need a site where code gets discussed/summarized/described (in small bites) by the science software community.
        The target audience would be graduate and undergraduate students in the sciences, and there would need to be
        mechanisms to keep the review positive and constructive, and not get pedantic or judgemental.  
        This could be tied to a software registry (such as the nascent GeoSoft project), or to language repositories like 
        R's CRAN repository, and could lead to the report card discussed in Action 2.
\end{itemize}

>>>>

Subgroup A consisted of ten participants who discussed papers surrounding training
and successful community software initiatives.
Subgroup B had six participants who discussed four papers. Because of the nature
of the papers, training emerged as a common theme. However the conversation was
wide-ranging, including incentives, reproducibility, and funding to promote
sustainability.  In the end, both groups discussed training, though from somewhat different
points of view and resulting in somewhat different suggested actions.

@@

Three main actions were identified by Subgroup A that would be of interest to participants and that
would benefit the research community.  These focused on the desire to amplify the 
current community efforts in training and software engineering by connecting the current training initiatives
(e.g., software/data carpentry, rOpenSci, OSS) that are of different durations at appropriate stages. Generally it 
was felt that a modicum of interaction between short term workshops (SwC, DC), medium term trainings (e.g., OSS),
and longer-term courses (e.g., BIDS Data Science) would benefit from coordinating curriculum,
discussing and aligning prerequisites, and coordinating timing of courses.  A training
coordination effort would go a long ways towards amplifying the value of the individual
efforts and make them all more effective.

\begin{itemize}
    \item Action 1: Create a roadmap of research software training initiatives. Such a roadmap would
        provide a taxonomy of training opportunities: what they deliver, and what attendees need to know 
        going in (prerequisites). It would also show which recommended roadmap actions will have a clear 
        and immediate payoff, and which will have long-term payoffs.  The training roadmap would emphasize 
        time savings and efficiency gains to be had from each training.

    \item Action 2: Build a report card characterizing use of best practices in scientific software. Generally,
        people felt that researchers would be very willing to migrate practices if they could identify
        where they needed to improve. Such a report card would ask simple questions to characterize use of best
        practices in science software, and could be structured similarly to Joel Spolsky's Software Maturity
        questionnaire (also known as \href{http://www.joelonsoftware.com/articles/fog0000000043.html}{the Joel Test}~\cite{joel-test}).  The survey would create a 
        report card that shows areas where a project could improve, and then link those areas to specific 
        training offerings from the training taxonomy from Action 1.

    \item Action 3: Create a science software review forum.  It was generally acknowledged that a little code review
        can have a tremendous impact on the quality of software in a project, but that sites for science code
        review are lacking.  While people can ask questions on the Stack Exchange sites, they are generally not open
        to questions of style, approach, or appropriateness, as they try to avoid subjective commentary.   Instead, 
        we need a site where code gets discussed/summarized/described (in small bites) by the science software community.
        The target audience would be graduate and undergraduate students in the sciences, and there would need to be
        mechanisms to keep the review positive and constructive, and not get pedantic or judgmental.  
        This could be tied to a software registry (such as the nascent GeoSoft project), or to language repositories like 
        R's CRAN repository, and could lead to the report card discussed in Action 2.
\end{itemize}
